% ---- ETD Document Class and Useful Packages ---- %
\documentclass{ucetd}
\usepackage{subcaption,epsfig,amsfonts}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mysty}

%% Use these commands to set biographic information for the title page:
\title{Data driven modeling of the low-Atwood single-mode Rayleigh-Taylor instability}
\author{Maxwell Hutchinson}
\department{Physics}
\division{Physical Sciences}
\degree{Doctor of Philosophy}
\date{December 9th, 2016}

%% Use these commands to set a dedication and epigraph text
\dedication{Dedicated to Tracey Ziev}
\epigraph{Epigraph Text}


\begin{document}
%% Basic setup commands
% If you don't want a title page comment out the next line and uncomment the line after it:
\maketitle
%\omittitle

% These lines can be commented out to disable the copyright/dedication/epigraph pages
\makecopyright
\makededication
\makeepigraph


%% Make the various tables of contents
\tableofcontents
\listoffigures
\listoftables

\acknowledgments
% Enter Acknowledgements here

I would like to thank Robert Rosner for letting me chart my course through this thesis and my doctoral research more broadly.
I thank Aleksandr Obabko and Paul Fischer for supporting Nek5000 and working with me through the numerics.
I also thank Oana Marin and Michel Schanen for their excitement and Elia Merzari and Ron Rahaman for their support.
I thank Alexander Heinecke and Scott Parker for their help tuning the code to run on supercomputers.
I thank Elizabeth Hicks for helpful comments and advice.

For computer time, this research partially used the resources of the
Supercomputing Laboratory at King Abdullah University of Science \& Technology
 (KAUST) in Thuwal, Saudi Arabia.
This research used resources of the Argonne Leadership Computing Facility, which
is a DOE Office of Science User Facility supported under Contract DE-AC02-06CH11357.
I acknowledge support from the Department of Energy Computational Science graduate fellowship.

\abstract
% Enter Abstract here

The Rayleigh-Taylor instability is one of the most common and well studied phenomena in fluid dynamics.
Despite research dating to
the late 19th century, the non-linear dynamics
of the interfacial instability are still not fully understood, particularly in the case when the two fluids have nearly the same density.
It was recently demonstrated in this, the low-Atwood regime, that the idealized single-mode problem departs from established potential flow models in the form of a re-acceleration beyond the predicted terminal interface velocity.
This thesis is an attempt to model that re-acceleration and, more broadly, the late time dynamics of the single-mode low-Atwood Rayleigh-Taylor instability.

The approach taken here is based on buoyancy-drag models, which express a force balance between buoyancy and parasitic drag.
The dynamical buoyancy-drag model is supplemented with a mixing model that dilutes the buoyant force over time.
These models are written deliberately generally, with 8 unique coefficients.
Three of these coefficients are solved for by equating the early time behavior with that of well established linear theories.
The remaining 5 coefficients are estimated by relating them to drag coefficients, friction factors, and geometric ratios in the interface shape.

To evaluate the model and compute the 5 unknown coefficients more precisely, a set of direct numerical simulations are performed over the relevant parameter space.
These simulations are first validated against experimental data.
Then they are shown to converge and their resolutions are chosen such as to minimize computational cost given the accuracy scale of the model.
The 5 coefficients are fit to the resulting data set, and the model achieves better than 2\% error in the bubble height and 4\% error in the volume of mixed fluid.
Three coefficients are nominally independent of the parameterization of the problem, while two are shown to vary with the Rayleigh number and the diffusivity.


\mainmatter
% Main body of text follows

\input{Chapters/Introduction} % Introduction

\input{Chapters/Background} % Introduction

\input{Chapters/Standalone}

\input{Chapters/Model}

\input{Chapters/Validation}

\input{Chapters/Conclusions} % Conclusion

% Format a LaTeX bibliography
\makebibliography

% Figures and tables, if you decide to leave them to the end
%\input{figure}
%\input{table}

\end{document}


