\newcommand{\pluseq}{\mathrel{+}=}

% make bold shape available for listings
\renewcommand{\ttdefault}{pcr}
\lstset{language=[90]Fortran,
  basicstyle=\ttfamily\lst@ifdisplaystyle\scriptsize\fi,
  keywordstyle=\bfseries,
  comment=[l]{!\ },
  escapechar=@,
%  numbers=left,
  tabsize=4,
}


\chapter{Efficiency of High Order Spectral Element Methods on Petascale Architectures}

\makeatletter
\newcommand{\chapterauthor}[1]{%
\begin{center}
 {\parindent0pt\vspace*{-25pt}%
  \linespread{1.1}\large\scshape#1%
  \par\nobreak\vspace*{35pt}}
\end{center}
  \@afterheading%
}
\makeatother

\vspace{20pt}

\chapterauthor{Maxwell Hutchinson, Alexander Heinecke, Hans Pabst, Greg Henry, Matteo Parsani, and David Keyes}

% a short form should be given in case it is too long for the running head
%\titlerunning{Lecture Notes in Computer Science: Authors' Instructions}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
%\author{Maxwell Hutchinson\inst{1} \and Alexander Heinecke\inst{2}  \and Hans Pabst\inst{3}  
%\and Greg Henry\inst{4}
%\and Matteo Parsani\inst{5}
%\and David Keyes\inst{5}
%}

%\institute{
%Department of Physics, University of Chicago, 5720 S. Ellis Ave, Chicago IL, 60637, USA
%Department of Physics, University of Chicago, Chicago IL, USA
%\and
%Intel Corporation, 2200 Mission College Blvd., Santa Clara CA, 95054, USA
%Intel Corporation, Santa Clara CA, USA
%\and
%Intel Semiconductor AG, Badenerstrasse 549, 8048 Zurich, Switzerland
%Intel Semiconductor AG, Zurich, Switzerland
%\and
%Intel Corporation, 2111 NE 25th Avenue, Hillsboro OR, 97124, USA
%Intel Corporation, Hillsboro OR, USA
%\and
%Extreme Computing Research Center, KAUST, Thuwal, 23955, KSA
%}


%\maketitle


\section{Abstract}
High order methods for the solution of PDEs expose a trade-off between computational cost and accuracy on a per degree of freedom basis.
In many cases, the cost increases due to higher arithmetic intensity while affecting data movement minimally.
As architectures tend towards wider vector instructions and expect higher arithmetic intensities, the best order for a particular simulation may change.

This study highlights preferred orders by identifying the high order efficiency frontier of the spectral element method implemented in Nek5000 and NekBox: the set of orders and meshes that minimize computational cost at fixed accuracy.
First, we extract Nek's order-dependent computational kernels and demonstrate exceptional hardware utilization by hardware-aware implementations.
Then, we perform production-scale calculations of the nonlinear single mode Rayleigh-Taylor instability on BlueGene/Q and Cray XC40-based supercomputers to highlight the influence of the architecture.
Accuracy is defined with respect to physical observables, and computational costs are measured by the core-hour charge of the entire application.
The total number of grid points needed to achieve a given accuracy is reduced by increasing the polynomial order.
On the XC40 and BlueGene/Q, polynomial orders as high as 31 and 15 come at no marginal cost per timestep, respectively.
Taken together, these observations lead to a strong preference for high order discretizations that use fewer degrees of freedom.
From a performance point of view, we demonstrate up to 60\% full application bandwidth utilization at scale and achieve $\approx$ 1 PFlop/s of compute performance in Nek's most flop-intense methods.

\section{Introduction}
\label{sec:introduction}
\input{Chapters/convergence/introduction}

\section{Nek's Computational Core}
\label{sec:math}
\input{Chapters/convergence/math}

\section{Kernel Analysis and Optimization}
\label{sec:implementation}
\input{Chapters/convergence/implementation}

\section{Scenarios and Performance}
\label{sec:benchmarks}
\input{Chapters/convergence/benchmarks}

\section{Conclusion}
\label{sec:conclusion}
\input{Chapters/convergence/conclusion}


{ \scriptsize

\noindent Optimization Notice: Software and workloads used in
performance tests may have been optimized for performance only on
Intel microprocessors.  Performance tests, such as SYSmark and
MobileMark, are measured using specific computer systems,
components, software, operations and functions.  Any change to any
of those factors may cause the results to vary.  You should
consult other information and performance tests to assist you in
fully evaluating your contemplated purchases, including the
performance of that product when combined with other products.
For more information go to http://www.intel.com/performance.

\noindent Intel, Xeon, and Intel Xeon
Phi are trademarks of Intel Corporation in the U.S. and/or other
countries.

}
