\chapter{Standalone papers}

This thesis contains three standalone papers.
They are logically ordered from the top down.

The first paper contains a new buoyancy-drag model developed to describe the late time behavior in the low-Atwood, moderate Grashof case.
It contains model coefficients that are fit to a data set of direct numerical simulations.
This single author paper will be submitted to a peer reviewed journal.

The second paper validates those direct numerical simulations against experimental data from Wilkinson and Jacobs~\cite{Wilkinson2007}.
This establishes that the simulations contain the physical processes responsible for re-acceleration and other unexplained late-time phenomena.
The paper takes advantage of the generality of numerical data to explore features of the flow that were not available to the original experiment, namely pressure driven secondary flows in the mid-plane and the interaction of the bubbles with one another.
This single author paper has been submitted to a peer reviewed journal.

The third paper contains a performance and convergence study of the numerical method and simulation software with the single mode Rayleigh-Taylor problem as a benchmark.
The NekBox code, which was specialized specifically to this project, is shown to be an efficient tool for performing these calculations.
Furthermore, the resolution is selected such that the simulation error is an order smaller than the expected model error, which ensures that the flow is sufficiently but not overly resolved.
Maxwell Hutchinson authored all of sections 2 and 4 and the majority of sections 1 and 5.
This paper has been peer reviewed and published in the proceedings of the International Conference on High Performance Computing (ISC).
