

\chapter{Direct numerical simulation of single mode three-dimensional Rayleigh-Taylor experiments}

\section{Abstract}
The single-mode Rayleigh-Taylor instability (smRTI) is well defined, poorly understood, and applicable to many fluid flows directly and through its relationship to multi-mode Rayleigh-Taylor models.
This study reproduces three low-Atwood smRTI experimental runs (Wilkinson and Jacobs, 2007) in
a specialized version of the Nek5000 spectral element code.
The simulations use the initial amplitude, wavelength, acceleration, Atwood number, and viscosity from 
the three specific experiments and impose no-slip and no-flux boundaries on the velocity and scalar, respectively.
The simulations are shown to reproduce the linear, saturation, stagnation, and re-acceleration phases of the smRTI
seen in the experiments.
Additionally, access to the full velocity and scalar fields demonstrates three different finite size effects: 
wall drag, wall lift, and a long wavelength mode along the diagonal.
One of the simulations is extended by a factor of two in the vertical direction and the resulting late-time dynamics
reach Froude numbers around 1.8, higher than previously reported.
Finally, inspection of the span-wise flow reveals secondary flows of the first kind that transport the scalar 
from the bubble-spike interfaces into the bubble and spike centers.
The agreement between simulations and experiments inspires confidence in the spectral element method for studying
the Rayleigh-Taylor instability.


\input{Chapters/validation/intro}

In \sref{methods}, we review the spectral element method and describe the numerical parameters of the simulations.
In \sref{results}, we compare the numerical and experimental results, extend the domain in extent and time to reach higher aspect ratios, and introduce new behavior in the span-wise flow.
In \sref{concs}, we discuss the validity of our methods for simulating Rayleigh-Taylor flows, the limits of wall-bounded single mode experiments, and the implications of secondary flows to mixing.
%\aref{data} provides access to the source, input, and output files for the numerical methods.

\input{Chapters/validation/methods}

\input{Chapters/validation/results}

\input{Chapters/validation/concs}
